---
title: "BPDA_PUMS"
author: "Yirong Yuan"
date: "2022-11-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install tidycensus package and get Census API Key from
http://api.census.gov/data/key_signup.html. 
Please read the instruction 
https://walker-data.com/tidycensus/articles/basic-usage.html
```{r}
library(tidycensus)
library(magrittr)
library(tidyverse)
library(dplyr)
Sys.getenv("CENSUS_API_KEY")
```


##Qurerying Data From Census
##PartI Countries' Population Datasets
This example only acquire the data in 2018. When you want to do other years, just 
change year = 2019/2017/.../2005.

```{r}
v1 = c("POBP","SEX","AGEP","MAR","YOEP","CIT","ENG","SCHL","COW","INDP","OCCP","POVPIP","WAGP")
v2<-c('AL','AK','AZ','AR','CA','CO','CT','DE','DC','FL','GA','HI','ID','IL','IN','IA','KS','KY','LA','ME','MD','MA','MI','MN','MS','MO','MT','NE','NV','NH','NJ','NM','NY','NC','ND','OH','OK','OR','PA','RI','SC','SD','TN','TX','UT','VT','VA','WA','WV','WI','WY')
pop_2018 <- get_pums(
  variables = v1,
  variables_filter = list(CIT=4:5),
  state = v2,
  survey = 'acs1',
  year = 2018
)
```

##separat different countries' data

```{r}
brazil_pr_2018<- grep("362", 
                   pop_2018$POBP, 
                   ignore.case = T)

brazil_pop_2018 <- pop_2018 %>% slice(brazil_pr_2018, preserve = FALSE)


argentina_pr_2018<- grep("360", 
                      pop_2018$POBP, 
                      ignore.case = T)

argentina_pop_2018 <- pop_2018 %>% slice(argentina_pr_2018, preserve = FALSE)

bolivia_pr_2018<- grep("361", 
                         pop_2018$POBP, 
                         ignore.case = T)

bolivia_pop_2018 <- pop_2018 %>% slice(bolivia_pr_2018, preserve = FALSE)

chile_pr_2018<- grep("363", 
                         pop_2018$POBP, 
                         ignore.case = T)

chile_pop_2018 <- pop_2018 %>% slice(chile_pr_2018, preserve = FALSE)

colombia_pr_2018<- grep("364", 
                         pop_2018$POBP, 
                         ignore.case = T)

colombia_pop_2018 <- pop_2018 %>% slice(colombia_pr_2018, preserve = FALSE)

ecuador_pr_2018<- grep("365", 
                         pop_2018$POBP, 
                         ignore.case = T)

ecuador_pop_2018 <- pop_2018 %>% slice(ecuador_pr_2018, preserve = FALSE)

french_guiana_pr_2018<- grep("367", 
                         pop_2018$POBP, 
                         ignore.case = T)

french_guiana_pop_2018 <- pop_2018 %>% slice(french_guiana_pr_2018, preserve = FALSE)

guyana_pr_2018<- grep("368", 
                             pop_2018$POBP, 
                             ignore.case = T)

guyana_pop_2018 <- pop_2018 %>% slice(guyana_pr_2018, preserve = FALSE)

paraguay_pr_2018<- grep("369", 
                             pop_2018$POBP, 
                             ignore.case = T)

paraguay_pop_2018 <- pop_2018 %>% slice(paraguay_pr_2018, preserve = FALSE)

peru_pr_2018<- grep("370", 
                     pop_2018$POBP, 
                     ignore.case = T)

peru_pop_2018 <- pop_2018 %>% slice(peru_pr_2018, preserve = FALSE)

suriname_pr_2018<- grep("371", 
                     pop_2018$POBP, 
                     ignore.case = T)

suriname_pop_2018 <- pop_2018 %>% slice(suriname_pr_2018, preserve = FALSE)

uruguay_pr_2018<- grep("372", 
                     pop_2018$POBP, 
                     ignore.case = T)

uruguay_pop_2018 <- pop_2018 %>% slice(uruguay_pr_2018, preserve = FALSE)

venezuela_pr_2018<- grep("373", 
                     pop_2018$POBP, 
                     ignore.case = T)

venezuela_pop_2018 <- pop_2018 %>% slice(venezuela_pr_2018, preserve = FALSE)


```
##standard data tidy for pop data
You can use this data to clean all other countires 

##clean population
```{r}
age <- brazil_pop_2018[,c("PWGTP","ST","AGEP")]
age$group <- "Population"
t14 <- age %>% group_by(ST,group) %>% summarise(n = sum(PWGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n)
t14<-column_to_rownames(t14, var = "group")
```


##Clean age

```{r}
age$group <- "age_0_to_4"
age[(age$AGEP < 9 & age$AGEP >= 5),"group"] <- "age_5_to_9"
age[(age$AGEP < 19 & age$AGEP >= 10),"group"] <- "age_10_to_19"
age[(age$AGEP < 24 & age$AGEP >= 20), "group"] <- "age_20_to_24"
age[(age$AGEP < 29 & age$AGEP >= 25), "group"] <- "age_25_to_29"
age[(age$AGEP < 34 & age$AGEP >= 30), "group"] <- "age_30_to_34"
age[(age$AGEP < 39 & age$AGEP >= 35), "group"] <- "age_35_to_39"
age[(age$AGEP < 44 & age$AGEP >= 40), "group"] <- "age_40_to_44"
age[(age$AGEP < 49 & age$AGEP >= 45), "group"] <- "age_45_to_49"
age[(age$AGEP < 54 & age$AGEP >= 50), "group"] <- "age_50_to_54"
age[(age$AGEP < 59 & age$AGEP >= 55), "group"] <- "age_55_to_59"
age[(age$AGEP < 64 & age$AGEP >= 60), "group"] <- "age_60_to_64"
age[(age$AGEP < 69 & age$AGEP >= 65), "group"] <- "age_65_to_69"
age[(age$AGEP < 74 & age$AGEP >= 70), "group"] <- "age_70_to_74"
age[(age$AGEP < 79 & age$AGEP >= 75), "group"] <- "age_75_to_79"
age[(age$AGEP < 84 & age$AGEP >= 80), "group"] <- "age_80_to_84"
age[(age$AGEP >= 85), "group"] <- "age_85_over"

t1 <- age %>% group_by(ST,group) %>% summarise(n = sum(PWGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n)
t1<-column_to_rownames(t1, var = "group")
```

##Clean gender
```{r}
sex <- brazil_pop_2018[,c("PWGTP","ST","SEX")]
sex$group <- "male"
sex[(sex$SEX=="2"),"group"] <- "female"

t2 <- sex %>% group_by(ST,group) %>% summarise(n = sum(PWGTP))%>% 
  pivot_wider(names_from = ST, values_from = n)
t2<-column_to_rownames(t2, var = "group")
```
## Clean Marital

```{r}
mar <- brazil_pop_2018[,c("PWGTP","ST","MAR")]
mar$group <- "Married"
mar[(mar$MAR=="2"),"group"] <- "Widowed"
mar[(mar$MAR=="3"),"group"] <- "Divorced"
mar[(mar$MAR=="4"),"group"] <- "Separated"
mar[(mar$MAR=="5"),"group"] <- "Never married or under 15 years old"
t3 <- mar %>% group_by(ST,group) %>% summarise(n = sum(PWGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n)
t3<-column_to_rownames(t3, var = "group")
```
##Clean Citizenship

```{r}
cit <- brazil_pop_2018[,c("PWGTP","ST","CIT")]
cit$group <- "US_citizen"
cit[(cit$CIT =="1" &cit$CIT =="2"& cit$CIT =="3" & cit$CIT =="4"),"group"] <- "US_citizen"
cit[(cit$CIT =="5"),"group"] <- "Non_US_citizen"
t4<-cit %>% group_by(ST,group) %>% count(group) %>% 
  pivot_wider(names_from = ST, values_from = n)
t4<-column_to_rownames(t4, var = "group")
```
##Clean Year of Entry

```{r}
yoe <- brazil_pop_2018[,c("PWGTP","ST", "YOEP")]
yoe$group <- "Entered US in 2000 or later"
yoe[(yoe$YOEP < 2000 ),"group"] <- " Entered US before 2000"

t5 <- yoe %>% group_by(ST,group) %>% count(group) %>% 
  pivot_wider(names_from = ST, values_from = n)
t5<-column_to_rownames(t5, var = "group")
```
##Clean Ability to speak English 

```{r}
eng <- brazil_pop_2018[,c("PWGTP","ST","ENG")]
eng$ENG[eng$ENG %in% c("b","1","2")] <- "Speaks_Englishty_Very_Well"
eng$ENG[eng$ENG %in% c("3","4")] <- "Speaks_English_Less_Than_Very_Well"
options(dplyr.summarise.inform = FALSE)
t6<- eng %>% group_by(ST,ENG) %>% summarise(n = sum(PWGTP))%>% 
  pivot_wider(names_from = ST, values_from = n)
t6<-column_to_rownames(t6, var = "ENG")
```
##Educational Attainment of population aged 25 or over

```{r}
edu <- brazil_pop_2018[,c("PWGTP","ST", "SCHL")]
edu$group <- "Less than high school"
edu[(edu$SCHL == "16" | edu$SCHL == "17" ),"group"] <- "High School Graduate/GED or Alternative Credential"
edu[(edu$SCHL == "18" 
     | edu$SCHL == "19" 
     | edu$SCHL == "20" ),"group"] <- "Some College/Associate Degree"
edu[(edu$SCHL == "21" 
     | edu$SCHL == "22" 
     | edu$SCHL == "23" 
     | edu$SCHL == "24"),"group"] <- "Bachelors Degree or Higher"

t7 <- edu %>% group_by(ST,group) %>% summarise(n = sum(PWGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n)
t7<-column_to_rownames(t7, var = "group")
```

##Employment Type

```{r}
cow <- brazil_pop_2018[,c("PWGTP","ST","COW")]
cow <-filter(cow,!(COW=="b"))
cow$group <- ""
cow[(cow$COW == "9"),"group"] <- "Unemployment"
cow[(cow$COW == '1' | cow$COW == '2'),"group"] <- "Private Wage and Salary Workers"
cow[(cow$COW == '3' | cow$COW == '4' |cow$COW == '5'),"group"] <- "Government Workers"
cow[(cow$COW == "6" | cow$COW == "7"),"group"] <- "Self employed Not incorporated"
cow[(cow$COW == "8"),"group"] <- "Self employed Incorporated"
t8 <- cow %>% group_by(ST,group) %>% summarise(n = sum(PWGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n)
t8<-column_to_rownames(t8, var = "group")
```
##Employment by Industry data clean

```{r}
indp <- brazil_pop_2018[,c("PWGTP","ST","INDP")]

indp$INDP[indp$INDP %in% c("0170","0180","0190","0270" ,"0280" ,"0290")]<-
  "Employment by Industry: AGR"

indp$INDP[indp$INDP %in% c("0370","0380","0390" ,"0470","0490")] <- 
  "Employment by Industry: EXT"

indp$INDP[indp$INDP %in% c("0570","0580","0590","0670","0680","0690")] <-
            "Employment by Industry: UTL"

indp$INDP[indp$INDP %in% c("0770")] <- "Employment by Industry: CON"

indp$INDP[indp$INDP %in% c("1070","1080","1090","1170","1180","1190",
                           "1270","1280" ,"1290","1370","1380" ,"1390",
                           "1470" ,"1480" ,"1490" ,"1570" ,"1590" ,"1670",
                           "1691" ,"1770" ,"1790","1870" ,"1880" ,"1890",
                           "1990","2070","2090","2170" ,"2180" ,"2190" ,
                           "2270","2280","2290","2370","2380" ,"2390",
                           "2470" ,"2480","2490" ,"2570","2590" ,"2670",
                           "2680" ,"2690" ,"2770" ,"2780" ,"2790" ,"2870",
                           "2880" ,"2890","2970","2980" ,"2990" ,"3070",
                           "1990" ,"2070" ,"2090","3080" ,"3095","3170",
                           "3180","3291" ,"3365","3470","3490" ,"3570","3580",
                           "3590","3670","3680" ,"3690" ,"3770","3780",
                           "3790","3875","3895" ,"3960" ,"3970","3980" ,"3990")]<-
  "Employment by Industry: MFG"

indp$INDP[indp$INDP %in% c("4070","4080","4090","4170","4180","4195" ,
                           "4265" ,"4270","4280" ,"4290","4370" ,"4380" ,
                           "4390" ,"4470" ,"4480","4490","4560","4570",
                           "4580","4585","4590")]  <- 
  "Employment by Industry: WHL"
  


indp$INDP[indp$INDP %in% c("4670" ,"4680","4690" ,"4770","4780" ,"4795",
                           "4870" ,"4880" ,"4890" ,"4971","4972" ,"4980",
                           "4990" ,"5070","5080" ,"5090","5170" ,"5180",
                           "5190" ,"5275","5280","5295" ,"5370","5381",
                           "5391","5470" ,"5480","5490" ,"5570" ,"5580",
                           "5593","5670" ,"5680","5690" ,"5790")] <- 
  "Employment by Industry: RET"

indp$INDP[indp$INDP %in% c("6070","6080","6090","6170","6180","6190",
                           "6270","6280","6290","6370","6380" ,"6390")] <- 
  "Employment by Industry: TRN"


indp$INDP[indp$INDP %in% c("6470","6480","6490","6570","6590","6670","6672",
                           "6680","6690","6695","6770","6780" )] <- 
  "Employment by Industry: INF"

indp$INDP[indp$INDP %in% c("6870","6880" ,"6890","6970","6991","6992","7071",
                      "7072","7080","7181","7190")] <- 
  "Employment by Industry: FIN"

indp$INDP[indp$INDP %in% c("7270","7280","7290" ,"7370","7380" ,"7390","7460",
                      "7470" ,"7480" ,"7490","7570","7670","7680" ,"7690" ,
                      "7770","7780","7790")] <- "Employment by Industry: PRF"

indp$INDP[indp$INDP %in% c("7860","7870","7880" ,"7890")] <- 
  "Employment by Industry: EDU"

indp$INDP[indp$INDP %in% c("7970" ,"7980","7990","8070","8080" ,"8090","8170","
                      8180","8191","8192","8270","8290" )] <- 
  "Employment by Industry: MED"

indp$INDP[indp$INDP %in% c("8370" ,"8380","8390","8470")] <-
  "Employment by Industry: SCA"

indp$INDP[indp$INDP %in% c("8561","8562","8563","8564","8570","8580",
                           "8590" ,"8660","8670","8680","8690")] <- 
  "Employment by Industry: ENT"


indp$INDP[indp$INDP %in% c("8770","8780","8790","8870","8891","8970",
                           "8980","8990","9070","9080","9090","9160",
                           "9170","9180","9190","9290")] <- 
  "Employment by Industry: SRV"

indp$INDP[indp$INDP %in% c("9370","9380","9390","9470","9480","9490",
                           "9570","9590")] <- 
  "Employment by Industry: ADM"

indp$INDP[indp$INDP %in% c("9490","9570","9590","9670" ,"9680" ,"9690",
                           "9770","9780" ,"9790","9870")] <- 
  "Employment by Industry: MIL"

options(dplyr.summarise.inform = FALSE)
t9<- indp %>% group_by(ST,INDP) %>% summarise(n = sum(PWGTP))%>% 
  pivot_wider(names_from = ST, values_from = n)
t9 <-filter(t9, !(INDP %in% c("0169", "9920","3370","7580","3380","7590",
                                 "3390", "8180" )))
t9<-column_to_rownames(t9, var = "INDP")
```
##Poverty Status
```{r}
povpip <- brazil_pop_2018[,c("PWGTP","ST","POVPIP")]
povpip$group <- "Not Determined"
povpip[(povpip$POVPIP <= 100 & povpip$POVPIP >0),"group"] <- "Individuals Below Poverty"
povpip[(povpip$POVPIP >100),"group"] <-"Individuals Above Poverty"
t10<-povpip %>% group_by(ST,group) %>% summarise(n = sum(PWGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n)
t10<-column_to_rownames(t10, var = "group")
```
##Median Personal Earnings
```{r}
wagp <- brazil_pop_2018[,c("PWGTP","ST","WAGP")]
t12 <- filter(wagp,WAGP>0)
t11<-wagp%>%group_by(ST) %>% summarise(n = median(rep(WAGP, PWGTP)))%>%
  pivot_wider(names_from = ST, values_from = n)

t12 %<>% group_by(ST) %>%summarise(n = median(rep(WAGP, PWGTP))) %>%
  pivot_wider(names_from = ST, values_from =n)

t13<-bind_rows(t11,t12)
rownames(t13)<- c("Median Personal Earnings","Median Personal Earnings (excl 0 and neg)")

```
##combine all subsets
```{r}
fn_brazi_pop<-bind_rows(t14,t1,t2,t3,t4,t5,t6,t7,t8,t9,t10,t13)
fn_brazi_pop <- fn_brazi_pop %>% replace(is.na(.), 0)
##If you want to export the data, using the following code
#write.csv(fn_brazi_pop, "fn_brazi_pop.csv")
```


##PartII Household data
Countries' Household Datasets
```{r}
v1<-c("POBP","HHT","NPF","POVPIP","OCPIP","GRPIP","ACR","HINCP",
                "FINCP")
v2<-c('AL','AK','AZ','AR','CA','CO','CT','DE','DC','FL','GA','HI','ID','IL','IN','IA','KS','KY','LA','ME','MD','MA','MI','MN','MS','MO','MT','NE','NV','NH','NJ','NM','NY','NC','ND','OH','OK','OR','PA','RI','SC','SD','TN','TX','UT','VT','VA','WA','WV','WI','WY')
household_2018 <- get_pums(
  variables = v1,
  variables_filter = list(SPORDER = 1),
  state = v2,
  survey = "acs1",
  year = 2018
)

```



```{r}

brazil_hr_2018<- grep("362", 
                      household_2018$POBP, 
                      ignore.case = T)

brazil_hou_2018 <- household_2018 %>% slice(brazil_hr_2018, preserve = FALSE)


argentina_hr_2018<- grep("360", 
                         household_2018$POBP, 
                         ignore.case = T)

argentina_hou_2018 <- household_2018 %>% slice(argentina_hr_2018, preserve = FALSE)

bolivia_hr_2018<- grep("361", 
                       household_2018$POBP, 
                       ignore.case = T)

bolivia_hou_2018 <- household_2018%>% slice(bolivia_hr_2018, preserve = FALSE)

chile_hr_2018<- grep("363", 
                     household_2018$POBP, 
                     ignore.case = T)

chile_hou_2018 <- household_2018 %>% slice(chile_hr_2018, preserve = FALSE)

colombia_hr_2018<- grep("364", 
                        household_2018$POBP, 
                        ignore.case = T)

colombia_hou_2018 <- household_2018%>% slice(colombia_hr_2018, preserve = FALSE)

ecuador_hr_2018<- grep("365", 
                       household_2018$POBP, 
                       ignore.case = T)

ecuador_hou_2018 <- household_2018 %>% slice(ecuador_hr_2018, preserve = FALSE)

french_guiana_hr_2018<- grep("367", 
                             household_2018$POBP, 
                             ignore.case = T)

french_guiana_hou_2018 <- household_2018 %>% slice(french_guiana_hr_2018, preserve = FALSE)

guyana_hr_2018<- grep("368", 
                      household_2018$POBP, 
                      ignore.case = T)

guyana_hou_2018 <- household_2018 %>% slice(guyana_hr_2018, preserve = FALSE)

paraguay_hr_2018<- grep("369", 
                        household_2018$POBP, 
                        ignore.case = T)

paraguay_hou_2018 <-household_2018 %>% slice(paraguay_hr_2018, preserve = FALSE)

peru_hr_2018<- grep("370", 
                    household_2018$POBP, 
                    ignore.case = T)

peru_hou_2018 <- household_2018 %>% slice(peru_hr_2018, preserve = FALSE)

suriname_hr_2018<- grep("371", 
                        household_2018$POBP, 
                        ignore.case = T)

suriname_hou_2018 <- household_2018%>% slice(suriname_hr_2018, preserve = FALSE)

uruguay_hr_2018<- grep("372", 
                       household_2018$POBP, 
                       ignore.case = T)

uruguay_hou_2018 <- household_2018 %>% slice(uruguay_hr_2018, preserve = FALSE)

venezuela_hr_2018<- grep("373", 
                         household_2018$POBP, 
                         ignore.case = T)

venezuela_hou_2018 <- household_2018 %>% slice(venezuela_hr_2018, preserve = FALSE)






```

##Clean household data
##Clean householder 
```{r}
hht<- brazil_hou_2018[,c("WGTP","ST","HHT")]
hht<-filter(hht,!(HHT=="b"))
hht$group <- "Householder"
b1 <- hht %>% group_by(ST,group) %>% summarise(n = sum(WGTP)) %>% pivot_wider(names_from = ST, values_from = n) 
b1<-column_to_rownames(b1, var = "group")

hht1<-filter(hht, HHT%in% c("1","2","3"))
hht1[(hht1$HHT ==  "1" | hht1$HHT ==  "2" | hht1$HHT ==  "3"),"group"] <- "Family Household"
b2<- hht1 %>% group_by(ST,group) %>% summarise(n = sum(WGTP)) %>% pivot_wider(names_from = ST, values_from = n)
b2<-column_to_rownames(b2, var = "group")
```

##Clean Ownership
```{r}

ocpip <- brazil_hou_2018[,c("WGTP","ST","OCPIP")]
ocpip$group <-"Owner Occupied Unit"
b3<- ocpip %>% group_by(ST,group) %>% summarise(n = sum(WGTP))%>%
  pivot_wider(names_from = ST, values_from = n)
b3<-column_to_rownames(b3, var = "group")

grpip<- brazil_hou_2018[,c("WGTP","ST","GRPIP")]
grpip$group<-" Renter Occupied Unit"
b4<- grpip %>% group_by(ST,group) %>% summarise(n = sum(WGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n)
b4<-column_to_rownames(b4, var = "group")
```

##clean Family Poverty
```{r}
povpip2 <- brazil_hou_2018[,c("WGTP","ST","POVPIP")]
povpip2$group <- ""
povpip2[(povpip2$POVPIP <= 100 ),"group"] <- " Families Below Poverty"
povpip2[(povpip2$POVPIP >100),"group"] <-"Families Above Poverty"
b5<-povpip2 %>% group_by(ST,group) %>% summarise(n = sum(WGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n)
b5<-column_to_rownames(b5, var = "group")

```
##Clean Monthly Owner Costs

```{r}
ocpip <- brazil_hou_2018[,c("WGTP","ST","OCPIP")]
ocpip$group <- "Monthly Owner Costs: Under 20% of HH Income"

ocpip[(ocpip$OCPIP < 25 & ocpip$OCPIP >= 20),"group"] <- 
  "Monthly Owner Costs: 20% to 25% of HH Income"
ocpip[(ocpip$OCPIP < 30 & ocpip$OCPIP >= 25),"group"] <- 
  "Monthly Owner Costs: 25% to 30% of HH Income"
ocpip[(ocpip$OCPIP < 35 & ocpip$OCPIP >= 30),"group"] <- 
  "Monthly Owner Costs: 30% to 35% of HH Income"
ocpip[(ocpip$OCPIP >= 35),"group"] <- 
  "Monthly Owner Costs: 35%+ of HH Income"
b6 <- ocpip %>% group_by(ST,group) %>% summarise(n = sum(WGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n) 
b6<-column_to_rownames(b6, var = "group")
```
 
##Clean Monthly Gross Rent
```{r}
grpip <- brazil_hou_2018[,c("WGTP","ST","GRPIP")]
grpip$group <- "Monthly Gross Rent: Under 15% of HH Income"

grpip[(grpip$GRPIP < 20 & grpip$GRPIP >= 15),"group"] <- 
  "Monthly Gross Rent: 15% to 20% of HH Income"
grpip[(grpip$GRPIP < 25 & grpip$GRPIP >= 20),"group"] <- 
  "Monthly Gross Rent: 20% to 25% of HH Income"
grpip[(grpip$GRPIP < 30 & grpip$GRPIP >= 25),"group"] <-
  "Monthly Gross Rent: 25% to 30% of HH Income"
grpip[(grpip$GRPIP < 35 &grpip$GRPIP >= 30),"group"] <- 
  "Monthly Gross Rent: 30% to 35% of HH Income"
grpip[(grpip$GRPIP >= 35),"group"] <- "Monthly Gross Rent: 35%+ of HH Income"

b7 <- grpip %>% group_by(ST,group) %>% summarise(n = sum(WGTP))%>% 
  pivot_wider(names_from = ST, values_from = n) 
b7<-column_to_rownames(b7, var = "group")
```

##clean Overcrowding data

```{r}
acr <- brazil_hou_2018[,c("WGTP","ST","ACR")]
acr <- filter(acr, !(ACR=="b"))
acr$group <- " Crowded Units"
acr[(acr$ACR== "2" | acr$ACR== "3"),"group"] <- 
  " Not Crowded Units"
b8 <- acr %>% group_by(ST,group) %>% summarise(n = sum(WGTP)) %>% 
  pivot_wider(names_from = ST, values_from = n) 
b8<-column_to_rownames(b8, var = "group")
```
##Clean Median Household Income

```{r}
hincp <- brazil_hou_2018[,c("WGTP","ST","HINCP")]

b9<- hincp %>% group_by(ST) %>%summarise(n = median(rep(HINCP, WGTP)) )%>%
  pivot_wider(names_from = ST, values_from = n)
rownames(b9) <- " Median Household Income"
```

##Median Family Income
```{r}
fincp <- brazil_hou_2018[,c("WGTP","ST","FINCP")]

b10<- fincp %>% group_by(ST) %>%summarise(n = median(rep(FINCP,WGTP))) %>%
  pivot_wider(names_from = ST, values_from = n)
rownames(b10) <- " Median Family Income"
```

##combine all subsets

```{r}
fn_brazi_hou<-bind_rows(b1,b3,b4,b2,b5,b6,b7,b8,b9,b10)
fn_brazi_hou <- fn_brazi_hou %>% replace(is.na(.), 0)
##If you want to export the data, using the following code
#write.csv(fn_brazi_pop, "fn_brazi_pop.csv")
```

